Using structural knowledge to correct RNA pyrosequencing errors in non-coding RNAs

### 2013-05-22 JCB/main_RNAPyro_JCB.tex
Initial commit JCB extended version
Started restructuring figures
New inside/outside figures
Almost there on outside...

\documentclass{llncs}
%\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage[colorinlistoftodos]{todonotes}

@@

\usepackage{tikz}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{relsize}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations}
\usetikzlibrary{snakes}
\pgfdeclarelayer{background}
\pgfdeclarelayer{backbackground}
\pgfdeclarelayer{foreground}
\pgfsetlayers{backbackground,background,main,foreground}

\newcommand{\ourprog}{\texttt{RNA-MoIP}\xspace}

\usepackage[applemac]{inputenc} %for the encoding

\newcommand{\RNAmutants}{\texttt{RNAmutants}\xspace}
\newcommand{\RNApyro}{\texttt{RNApyro}\xspace}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\farna}{\texttt{FARNA}\xspace}
\newcommand{\mcfoldmcsym}{\texttt{MC-Pipeline}\xspace}
\newcommand{\mcfold}{\texttt{MC-Fold}\xspace}
\newcommand{\mcsym}{\texttt{MC-Sym}\xspace}
\newcommand{\nast}{\texttt{NAST}\xspace}
\newcommand{\ifoldrna}{\texttt{iFoldRNA}\xspace}
\newcommand{\rnafold}{\texttt{RNAfold}\xspace}
\newcommand{\rnasubopt}{\texttt{RNAsubopt}\xspace}
\newcommand{\rnawolf}{\texttt{RNAwolf}\xspace}
\newcommand{\rnastructure}{\texttt{RNAstructure}\xspace}
\newcommand{\contrafold}{\texttt{contrafold}\xspace}
\newcommand{\unafold}{\texttt{unafold}\xspace}
\newcommand{\rnadd}{\texttt{RNA2D3D}\xspace}
\newcommand{\assemble}{\texttt{assemble}\xspace}
\newcommand{\fred}{\texttt{FR3D}\xspace}
\newcommand{\rnajunction}{\texttt{RNAjunction}\xspace}
\newcommand{\rnamotif}{\texttt{RNAmotif}\xspace}
\newcommand{\treefolder}{\texttt{TreeFolder}\xspace}
\newcommand{\barnacle}{\texttt{BARNACLE}\xspace}
\newcommand{\contextfold}{\texttt{contextfold}\xspace}
%\newcommand{\citep}{\cite}
\newcommand{\Z}[3]{\mathcal{Z}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\Y}[3]{\mathcal{Y}_{\substack{(#1)\\ [#3]}}^{#2}}
\newcommand{\B}{\mathcal{B}}
\newcommand{\Kron}{\delta}
\newcommand{\ub}{\bullet}

\newcommand{\Ab}{{\sf{A}}}
\newcommand{\Cb}{{\sf{C}}}
\newcommand{\Gb}{{\sf{G}}}
\newcommand{\Ub}{{\sf{U}}}

\newcommand{\SpaceCheating}{\vspace{-0em}}

%\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
\title{A Linear Inside-Outside Algorithm for Correcting Sequencing Errors in Structured RNAs}
\author{Vladimir Reinharz\inst{1}, Yann Ponty\inst{2}$^{*}$ \and J\'er\^{o}me Waldisp\"{u}hl\inst{1}$^*$}
\date{}
\institute{School of Computer Science, McGill University, Montreal, Canada.
	\and  Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.
	 \\\email{jeromew@cs.mcgill.ca}, \email{yann.ponty@lix.polytechnique.fr}}

@@

\begin{document}
\maketitle
%\section{}
%\subsection{}
\begin{abstract}
Analysis of the sequence-structure relationship in RNA molecules are essential to evolutionary studies but also to 
concrete applications such as error-correction methodologies in sequencing technologies. The prohibitive sizes of the
mutational and conformational landscapes combined with the volume of data to proceed require efficient algorithms 
to compute sequence-structure properties. More specifically, here we aim to calculate which mutations increase the most the 
likelihood of a sequence to a given structure and RNA family.\\
In this paper, we introduce \RNApyro, an efficient linear-time and space inside-outside algorithm that computes exact mutational
probabilities under secondary structure and evolutionary constraints given as a multiple sequence alignment with a consensus structure.
We develop a scoring scheme combining classical stacking base pair energies to novel isostericity scales, and apply our techniques
to correct point-wise errors in 5s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
tools in the NGS error-correction pipeline.

@@

\input{introduction_RECOMB.tex}

\input{methods_RECOMB.tex}

\input{results_RECOMB.tex}

\input{conclusion_RECOMB.tex}

\input{acknowledgments_RECOMB.tex}
%\newpage

\bibliographystyle{splncs}
\bibliography{RNApyro}

>>>>

\documentclass{llncs}
\usepackage[top=1.5in, bottom=1.5in,right=1.5in,left=1.5in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
%\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
%\usepackage{jcompbio}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage[colorinlistoftodos,textsize=small]{todonotes}

@@

\usepackage{tikz}
\usepackage{amsmath}
\usepackage{ifthen}
\usepackage{relsize}
\usetikzlibrary{positioning}
\usetikzlibrary{matrix}
\usetikzlibrary{calc}
\usetikzlibrary{shapes}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{decorations}
\usetikzlibrary{snakes}
\usetikzlibrary{shadows}
\pgfdeclarelayer{background}
\pgfdeclarelayer{backbackground}
\pgfdeclarelayer{foreground}
\pgfsetlayers{backbackground,background,main,foreground}

\input{macros.tex}

%\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
\title{Using structural knowledge to correct RNA pyrosequencing errors in non-coding RNAs}
\author{Vladimir Reinharz\inst{1}, Yann Ponty\inst{2}$^{*}$ \and J\'er\^{o}me Waldisp\"{u}hl\inst{1}$^*$}
\date{}
\institute{School of Computer Science, McGill University, Montreal, Canada.
	\and  Laboratoire d'informatique, \'Ecole Polytechnique, Palaiseau, France.
	 \\\email{jeromew@cs.mcgill.ca}, \email{yann.ponty@lix.polytechnique.fr}}

@@

\begin{document}

\ShowTODO{\setcounter{tocdepth}{1}
\listoftodos} 
\maketitle
%\section{}
%\subsection{}
\begin{abstract}
Analysis of the sequence-structure relationship in RNA molecules are essential to evolutionary studies but also to 
concrete applications such as error-correction methodologies in sequencing technologies. The prohibitive sizes of the
mutational and conformational landscapes combined with the volume of data to proceed require efficient algorithms 
to compute sequence-structure properties. More specifically, here we aim to calculate which mutations increase the most the 
likelihood of a sequence to a given structure and RNA family.\\
In this paper, we introduce \RNApyro, an efficient linear-time and space inside-outside algorithm that computes exact mutational
probabilities under secondary structure and evolutionary constraints given as a multiple sequence alignment with a consensus structure.
We develop a scoring scheme combining classical stacking base pair energies to novel isostericity scales, and apply our techniques
to correct point-wise errors in 5s rRNA sequences. Our results suggest that \RNApyro is a promising algorithm to complement existing
tools in the NGS error-correction pipeline.

@@

\input{introduction_JCB.tex}

\input{methods_JCB.tex}

\input{results_JCB.tex}

\input{conclusion_JCB.tex}

\input{acknowledgments_JCB.tex}
%\newpage

\bibliographystyle{jcompbiol}
\bibliography{RNApyro}
